### 2016-07-22 grlp.tex
changes up to discussion

%!TEX root =  autocontgrlp.tex
\section{Generalized Reduced Linear Program}
The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by adding the extra constraint $J=\Phi r$ in \eqref{mdplp} with $\Phi \in \R^{n\times k}$ and introducing the new variables $r\in \R^k$. By substitution, this leads to
\begin{align}\label{alp}
\begin{split}
\min_{r\in \R^k}\, &c^\top \Phi r\\
\text{s.t.}\mb &\Phi r\geq T \Phi r,
\end{split}
\end{align}
where $J\geq TJ$ is a shorthand for the $nd$ constraints in \eqref{mdplp}. Unless specified otherwise we use $\tr$ to denote an arbitrary solution to the ALP, and we let $\tj=\Phi \tr$ to denote the corresponding approximate value function and $\tu$ to denote the greedy policy with respect to $\tj$.
The Generalized Reduced Linear Program is given as:
\begin{align}\label{grlp}
\begin{split}
\underset{r\in \N}{\min}\, &\, c^\top \Phi r,\\
\text{s.t.}\mb & \,W^\top E\Phi r\geq W^\top H \Phi r,
\end{split}
\end{align}
where $W \in \R^{nd\times m}_+$ is a matrix with all positive entries and $\N$ is an additional (compact) constraint set to ensure the boundedness of the solution.
In what follows, we denote the solution to the GRLP by $\hr$, the approximate value function by $\hj=\Phi \hr$ and use $\hu$ to denote the greedy policy with respect to $\hj$.

For brevity, we introduce our assumptions, whose meaning will be explained later.
\begin{assumption}\label{grlpassmp}
\begin{enumerate}[(i)]
\item \label{probdist} $c=(c(i),i=1,\ldots,n)\in \R^n$ is a positive probability distribution, i.e., $c(i)>0 \mb \forall i$ and $\sum_{i=1}^n c(i)=1$.
\item  \label{one} The first column of the feature matrix $\Phi$ (i.e., $\phi_1$) is $\one \in \R^n$. 
\item \label{lyap} $\psi\colon S \ra \R_+$ is a Lyapunov function for $P$
and is present in the column span of the feature matrix $\Phi$: For some $r_0\in \R^k$, $\Phi r_0 = \psi$. 
\item \label{nassmp} $\N\subset \R^k$ is compact and $\tr \in \N\subset\R^k$.
\item \label{wassump} $W \in \R^{nd\times m}_+$ is a full rank $nd\times m$ matrix (where $m\ll nd$), with all non-negative entries such that each of its column-sums equals one.
\item \label{ass:n4} The set $\N'$ is such that $\N' = \N + t r_0$ for any $t\in \R$, where $r_0\in \R^k$ such that $\Phi r_0 = \psi$.
\end{enumerate}
\end{assumption}
We note in passing that if \cref{grlpassmp}-\eqref{lyap} holds, it follows that for any $J\in \R^n$ and $t>0$,
\begin{align}
\label{eq:psilin}
T(J+ t \psi ) \le TJ + \beta_{\psi}\,t\,  \psi\,. %\,\, \quad (J\in \R^n,\, t>0)\,.
\end{align}

The GRLP introduces linear function approximation in both the primal and dual variables of the LP formulation. 
To understand this, we need to look at the Lagrangian of the ALP and GRLP in \eqref{lag} and \eqref{lag2} respectively, i.e., 
\begin{align}\label{lag}
\tilde{L}(r,\lambda)=c^\top \Phi r+\lambda^\top (T\Phi r-\Phi r), \\ \label{lag2}\hat{L}(r,q)=c^\top \Phi r+q^\top W^\top (T\Phi r-\Phi r).
\end{align}
Note that $ Wq\approx \lambda$ in \eqref{lag2} is linear function approximation of the Lagrange multipliers. Note that while the ALP employs LFA in its objective function (i.e., use of $\Phi r$), the GRLP employs linear approximation both in the objective function ($\Phi r$) as well as the constraints (use of $W$). Further, $W$ can be interpreted as the feature matrix that approximates the Lagrange multipliers as $\lambda\approx Wq$, where $\lambda \in \R^{nd}, r\in \R^m$. One can show \cite{dolgov} that the optimal Lagrange multipliers are the discounted number of visits to the ``state-action pairs'' under an optimal policy $u^*$, i.e., 
\begin{align}
\lambda^*(s,u^*(s))&=\big(c^\top(I-\alpha P_{u^*})^{-1}\big)(s)\nn\\
				&= \big(c^\top(I+\alpha P_{u^*}+\alpha^2 P_{u^*}^2+\ldots)\big)(s),\nn\\
			\lambda^*(s,a)&=0, \qquad \text{for all } a \neq u^*(s),\nn
\end{align}
where $P_{u^*}$ is the probability transition matrix under $u^*$ ($P_{u^*}(s,s') = P_{u^*(s)}(s,s')$, $s,s'\in S$). Even though we might not have the optimal policy $u^*$ in practice, the fact that $\lambda^*$ is a probability distribution and that it is a linear combination of $\{P_{u^*},P^2_{u^*},\ldots\}$ hints at the kind of features that might be useful for the $W$ matrix.

>>>>

%!TEX root =  autocontgrlp.tex
\section{Generalized Reduced Linear Program}
The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by introducing the new variables $r\in \R^k$ and adding the extra constraint $J=\Phi r$ in \eqref{mdplp} with $\Phi \in \R^{n\times k}$ \citep{SchSei85}. 
By substitution, this leads to
\begin{align}\label{alp}
\begin{split}
\min_{r\in \R^k}\, &c^\top \Phi r\,\,\,
\text{s.t.}\mb \Phi r\geq T \Phi r,
\end{split}
\end{align}
where $J\geq TJ$ is a shorthand for the $nd$ constraints in \eqref{mdplp}. Unless specified otherwise we use $\tr$ to denote an arbitrary solution to the ALP, and we let $\tj=\Phi \tr$ to denote the corresponding approximate value function and $\tu$ to denote the greedy policy w.r.t. $\tj$.
The Generalized Reduced Linear Program is given as:
\begin{align}\label{grlp}
\begin{split}
\underset{r\in \N}{\min}\, &\, c^\top \Phi r\,\,\,\,
\text{s.t.}\mb  \,W^\top E\Phi r\geq W^\top H \Phi r,
\end{split}
\end{align}
where $W \in \R^{nd\times m}_+$ is a matrix with all positive entries and $\N$ is an additional (compact) constraint set to ensure the boundedness of the solution.%
\footnote{The appendix explains how $\N$ can be chosen.}
In what follows, we denote the solution to the GRLP by $\hr$, the approximate value function by $\hj=\Phi \hr$ and use $\hu$ to denote the greedy policy w.r.t. $\hj$.

We assume that the following hold in the rest of the paper:
\begin{assumption}\label{grlpassmp}
\begin{enumerate}[(i)]
\item \label{probdist} $c=(c(i),i=1,\ldots,n)\in \R^n$ is a positive probability distribution, i.e., $c(i)>0 \mb \forall i$ and $\sum_{i=1}^n c(i)=1$.
\item  \label{one} The first column of the feature matrix $\Phi$ (i.e., $\phi_1$) is $\one \in \R^n$. 
\item \label{lyap} $\psi\colon S \ra \R_+$ is a Lyapunov function for $P$
and is present in the column span of the feature matrix $\Phi$: For some $r_0\in \R^k$, $\Phi r_0 = \psi$. 
\item \label{nassmp} $\N\subset \R^k$ is compact and $\tr \in \N\subset\R^k$.
\item \label{wassump} $W \in \R^{nd\times m}_+$ is a full rank $nd\times m$ matrix (where $m\ll nd$), with all non-negative entries such that each of its column-sums equals one.
\item \label{ass:n4} The set $\N'$ is such that $\N' = \N + t r_0$ for any $t\in \R$, where $r_0\in \R^k$ such that $\Phi r_0 = \psi$.
\end{enumerate}
\end{assumption}
We note in passing that if \cref{grlpassmp}-\eqref{lyap} holds, it follows that for any $J\in \R^n$ and $t>0$,
\begin{align}
\label{eq:psilin}
T(J+ t \psi ) \le TJ + \beta_{\psi}\,t\,  \psi\,. %\,\, \quad (J\in \R^n,\, t>0)\,.
\end{align}

The GRLP introduces linear function approximation in both the primal and dual variables of the LP formulation. 
To understand this, we need to look at the Lagrangian of the ALP and GRLP in 
\eqref{lag} and \eqref{lag2} respectively, i.e., 
\begin{align}\label{lag}
\tilde{L}(r,\lambda)=c^\top \Phi r+\lambda^\top (T\Phi r-\Phi r), \\ \label{lag2}\hat{L}(r,q)=c^\top \Phi r+q^\top W^\top (T\Phi r-\Phi r).
\end{align}
Thus, when $Wq = \lambda$, i.e., when $W$ is a set of basis functions that allow
a low dimensional linear representation of the dual variables $\lambda$,
the two problems are the same.
%Note that $ Wq\approx \lambda$ in \eqref{lag2} is linear function approximation of the Lagrange multipliers. 
Hence, while the ALP employs LFA in its objective function (i.e., use of $\Phi r$), the GRLP employs linear approximation both in the objective function ($\Phi r$) as well as the constraints (use of $W$). 
%Further, $W$ can be interpreted as the feature matrix that approximates the Lagrange multipliers as $\lambda\approx Wq$, where $\lambda \in \R^{nd}, r\in \R^m$. 
To get a sense of how $W$ should be chosen, recall that
the optimal Lagrange multipliers are the discounted number of visits to the ``state-action pairs'' under an optimal policy $u^*$, i.e., 
\begin{align}
\lambda^*(s,u^*(s))&=\big(c^\top(I-\alpha P_{u^*})^{-1}\big)(s)\nn\\
				&= \big(c^\top(I+\alpha P_{u^*}+\alpha^2 P_{u^*}^2+\ldots)\big)(s),\nn\\
			\lambda^*(s,a)&=0, \qquad \text{for all } a \neq u^*(s),\nn
\end{align}
where $P_{u^*}$ is the probability transition matrix under $u^*$ ($P_{u^*}(s,s') = P_{u^*(s)}(s,s')$, $s,s'\in S$) \cite{dolgov}. Even though we might not have the optimal policy $u^*$ in practice, the fact that $\lambda^*$ is a probability distribution and that it is a linear combination of $\{P_{u^*},P^2_{u^*},\ldots\}$ hints at the kind of features that might be useful for the $W$ matrix.

### 2016-07-22 intro.tex
changes up to discussion

%!TEX root =  autocontgrlp.tex
\section{Introduction}
Markov decision processes (MDPs) is a powerful mathematical framework to study optimal sequential decision making problems that arise in science and engineering. The so-called dynamic programming methods find an optimal policy by computing the optimal \emph{value-function} ($J^*$), a vector whose dimension is the number of states. MDPs with small number of states can be solved easily by conventional solution methods such as value-, or policy-iteration, or linear programming (LP) \cite{BertB}.

A practical way to tackle the issue of large number of states is to compute an approximate value function $\tilde{J}$ instead of $J^*$. Here, success depends on the quality of approximation, i.e., on the quantity $||J^*-\tilde{J}||$ for an appropriately chosen norm. The most widely used method for approximation is the linear function approximation (LFA), i.e., let $\tilde{J}=\Phi r^*$, where $\Phi$ is a feature matrix and $r^*$ is a weight vector to be computed. Dimensionality reduction is achieved by choosing $\Phi$ to have fewer columns in comparison to the number of states, making it possible to work with large MDPs where even storing one value per state would be impossible in the computer's main memory.

The \emph{approximate linear program} (ALP) \cite{ALP,CS,SALP,ALP-Bor} and its variants introduce linear function approximation in the linear programming formulation. A critical shortcoming of the ALP is that the number of constraints are of the order of the size of the state space, making the vanilla version of the ALP intractable. A way out is to employ choose a subset of constraints random and drop the rest, thereby formulating a \emph{reduced linear program} (RLP). The performance analysis of the RLP can be found in \cite{CS} and the RLP has also been shown to perform well in experiments \cite{ALP,CS,CST}. An alternative approach to handle the issue of large number of constraints is to employ function approximation in the dual variables of the ALP \cite{ALP-Bor,dolgov}, an approach that was empirically found to be useful. However, to this date, there exist no theoretical guarantees bounding the loss in performance resulting from such an approximation.

In this paper, we generalize the RLP to define a generalized reduced linear program (GRLP) which has a tractable number of constraints obtained as positive linear combinations of the original constraints of the ALP. 
The salient aspects of our contribution are listed below:
\begin{enumerate}
		\item We develop novel analytical machinery to relate $\hat{J}$, the solution to the GRLP, and the optimal value function $J^*$ by bounding the prediction error $||J^*-\hj||$ (\Cref{cmt2mn}). 
		\item We also bound the performance loss due to using the one-step greedy policy $\hu$ (greedy with respect to $\hj$) as a function of $||J^*-J_{\hu}||$ (Theorem~\ref{polthe}).
		\item Our analysis is based on two novel $\max$-norm contraction operators and our results hold \emph{deterministically}, as opposed to the results on RLP \cite{SALP,CS}, where the guarantees have a probabilistic nature.
		\item Our results on the GRLP are the first to theoretically justify linear function approximation of Langrangian (dual) variables underlying the constraints.
		\item A numerical example in controlled queues is provided to illustrate the theory.
\end{enumerate}

>>>>

%!TEX root =  autocontgrlp.tex
\section{Introduction}
Markov decision processes (MDPs) are a powerful mathematical framework to study optimal sequential decision making problems  arising in science and engineering. The so-called dynamic programming methods find an optimal policy by computing the optimal \emph{value-function} ($J^*$), a vector whose dimension is the number of states. MDPs with small number of states can be solved easily by conventional dynamic programming techniques, such as value-, or policy-iteration, or linear programming (LP) \cite{BertB}.

In this paper we consider the problem of using dynamic programming with MDPs with large state spaces.
A practical way to tackle the issue of large number of states is to compute an approximate value function $\tilde{J}$ instead of $J^*$. %Here, success depends on the quality of approximation, i.e., on the quantity $||J^*-\tilde{J}||$ for an appropriately chosen norm. 
Linear function approximation (LFA), i.e., letting $\tilde{J}=\Phi r^*$ where $\Phi$ is a so-called feature matrix and $r^*$ is a weight vector to be computed, is the most widely used method of approximation. Here, dimensionality reduction is achieved by choosing $\Phi$ to have fewer columns in comparison to the number of states, holding the promise of being able to work with MDPs regardless of the number of states.

The \emph{approximate linear program} (ALP) \cite{ALP,CS,SALP,ALP-Bor} and its variants introduce linear function approximation in the linear programming formulation. A critical shortcoming of vanilla ALP is that the number of constraints are of the order of the size of the state space, making this vanilla version intractable. A way out is to choose a subset of constraints at random and drop the rest, thereby formulating a \emph{reduced linear program} (RLP). The performance analysis of the RLP can be found in \cite{CS} and the RLP has also been shown to perform well in experiments \cite{ALP,CS,CST}. An alternative approach to handle the issue of large number of constraints is to employ function approximation in the dual variables of the ALP \cite{ALP-Bor,dolgov}, an approach that was also found useful in experiments. However, to this date, there exist no theoretical guarantees bounding the loss in performance resulting from such an approximation.

In this paper, we generalize the RLP to define a generalized reduced linear program (GRLP) which has a tractable number of constraints that are obtained as positive linear combinations of the original constraints.
The salient aspects of our contribution are listed below:
\begin{enumerate}
		\item We develop novel analytical machinery to relate $\hat{J}$, the solution to the GRLP, and the optimal value function $J^*$ by bounding the prediction error $||J^*-\hj||$ (\Cref{cmt2mn}). 
		\item We also bound the performance loss due to using the policy $\hu$ that is one-step greedy with respect to $\hj$ (Theorem~\ref{polthe}).
		\item Our analysis is based on two novel $\max$-norm contraction operators and our results hold \emph{deterministically}, as opposed to the results on RLP \cite{SALP,CS}, where the guarantees have a probabilistic nature.
		\item Our results on the GRLP are the first to theoretically analyze the use of linear function approximation of Langrangian (dual) variables underlying the constraints.
		\item A numerical example in controlled queues is provided to illustrate the theory.
\end{enumerate}

### 2016-07-22 mdp.tex
changes up to discussion

A stationary deterministic policy\footnote{For the scope of this paper, it suffices to restrict our attention to stationary deterministic policies.} (SDP) or simply a policy is a map $u\colon S\ra A$ that specifies for each state what action to select in that state.
Given an SDP $u$, the expected total discounted reward corresponding to starting at state $s$ and then choosing actions as dictated by $u$ for the states encountered is denoted by $J_u(s)$ and is defined by
\begin{align}
J_u(s)\stackrel{\Delta}{=}\E[\sum_{n=0}^\infty \alpha^n g_{a_n}(s_n)|s_0=s,a_n=u(s_n)\mbox{ }\forall n\geq 0],\nn
\end{align}
where $\alpha \in (0,1)$ is the so-called discount factor and $s_{n+1} \sim p_{a_n}(s_n,\cdot)$, $n\ge 0$. 
Here $J_u(s)$ is known as the value of the state $s$ under SDP $u$, and $J_u$ (as a map from $S$ to the reals)
is known as the value function corresponding to the SDP $u$.
The \emph{optimal policy} $u^*$ is one that in each state $s\in S$ achieves the best possible total expected discounted reward from that state. That is, $J^{u^*})(s) = J^*(s) \eqdef \max_{u\in U} J_u(s)$.\footnote{In our case an optimal (SDP) $u^*$ exists and is well defined \cite{BertB}.}, where $U$ is the set of all SDPs and $J^*$ is coined the \emph{optimal value function}.

Any optimal policy $u^*$ and value function $J^*$ obey the Bellman equation (BE): for all $ s \in S$, 
\begin{subequations}\label{bell}
\begin{align}
\label{bellval}J^*(s)&=\max_{ a\in A}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big),\\
\label{bellpol}u^*(s)&= \underset{a\in A}{\argmax}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big)\,.
\end{align}
\end{subequations}
Once $J^*$ is computed, $u^*$ can be obtained via \eqref{bellpol}. Thus computing the value function is at the heart of most solution methods to MDP.

For convenience, in what follows we identify elements of $\R^n$ with $\R^S$ in the natural way. In particular, we can thus treat $J_u$ and $J^*$ as vectors in $\R^n$. Similarly we identify $\R^{nd}$ with $\R^{S\times A}$.
We now give some definitions that will be useful later.
%The Bellman operator $T$ is defined using the model parameters of the MDP as follows:
\begin{definition}
Let $c,\rho,\chi:S \to \R_+$ be positive valued functions, where $\R_+$ denotes the set of strictly positive reals. Then for $J\in \R^n$, $a\in A$ and $s\in S$, 
define
\begin{enumerate}[(i)]
\item The Bellman operator $T\colon \R^n \ra \R^n$ as $(TJ)(s)=\max_{a \in A}\big(g_a(s)+\alpha \sum_{s'} p_a(s,s')J(s')\big).
$
\item \label{bellactval} The Bellman operator (of action values) $H: \R^n \to \R^{nd}$ for state-action values as $HJ=[ H_1 J,\cdots,H_d J]^\top\in \R^{nd},$ where $(H_a J)(s)= g_a(s)+\alpha \sum_{s'}p_a(s,s') J(s')$.
\item The weighted $L_1$-norms $\norm{\cdot}_{1,c}$ 
and 
the weighted $L_\infty$-norms  $\norm{\cdot}_{\infty,\rho}$ as $
||J||_{1,c}=\sum_{s \in S} c(s)|J(s)|, 
||J||_{\infty,\rho}=\max_{s \in S} \frac{|J(s)|}{\rho(s)}$.
\item The discounted maximal inflation of $\chi$ due to $P = (p_a)_{a\in A}$ as $\beta_{\chi}=\max_{s \in S} \frac{\underset{a \in A}{\max}\big(\alpha\sum_{s'}p_a(s,s')\chi(s')\big)}{\chi(s)}$.
\item The function $\chi:S\to\R_+$ above to be a \emph{Lyapunov} function for $P = (p_a)_{a\in A}$ if $\beta_{\chi}<1$.
\item $E$ to be the $nd\times n$ matrix given by $E=[I,\ldots,I]^\top$, i.e., $E$ is obtained by stacking $d$ identical $n\times n$ identity matrices one over the other.
\item A policy $\tu$ is said to be greedy with respect to some function $\tj:S \to \R$ if \begin{align*}\tu\eqdef\underset{a\in A}{\argmax}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')\tj(s')\big).\end{align*}
\end{enumerate}
\end{definition}
We now state without proof the most important properties of the Bellman operator.
The proofs are immediate from the definitions, but can also be found in \cite{BertB}.
First, we introduce some extra notation:
For $J_1,J_2\in \R^n$, we write $J_1\le J_2$ if $J_1(s)\le J_2(s)$ holds for all $s\in S$.
We use $\one \in \R^n$ to denote a vector with all entries $1$.
The maximum norm $\norm{\cdot}_{\infty}$ is defined by $ \norm{v}_{\infty} = \max_{s\in S} |v(s)|$.
\begin{lemma}\label{monotone}
$T$ is a monotone map, i.e., given $J_1,J_2 \in \R^n$ such that $J_1\leq J_2$, we have $T J_1\leq T J_2$. 
\end{lemma}
\begin{lemma}\label{shift}
Given $J\in \R^n$ and $t \in \R$, we have
\begin{align}\label{eq:shift}
T(J+t\one)=TJ+\alpha t\one.
\end{align}
\end{lemma}
\begin{lemma}\label{maxnorm}
If $T: \R^n \to \R^n$ is any operator that is monotonous and satisfies~\eqref{eq:shift} then 
$T$ is a $\max$-norm contraction operator with contraction factor $\alpha \in (0,1)$, i.e., given $J_1, J_2 \in \mathbf{R}^n$,
\begin{align}
||TJ_1-TJ_2||_\infty\leq \alpha ||J_1-J_2||_\infty.
\end{align}
\end{lemma}
\begin{lemma}\label{uniquesol} 
$J^*$ is a unique fixed point of $T$, i.e., $J^*=TJ^*$.
\end{lemma}
\begin{corollary}
If $J\in \R^n$ is such that $J\geq TJ$ then $J\geq TJ^2\geq \ldots \geq J^*$.
\end{corollary}
Though  \cref{monotone,shift,maxnorm} are stated for the Bellman operator $T$, the results also hold for $H$ as well.

\subsection{The Linear Programming Formulation}
As it is well known, the optimal value function $J^*$ can be obtained by solving the following linear program:
\begin{align}\label{mdplp}
\min_{J\in \R^n}\, &c^\top J\\
\text{s.t.}\mb &J(s)\geq g_a(s)+\alpha\sum_{s'}p_a(s,s')J(s'),\nn\\
&\forall s \in S, a \in A,
\end{align}
where $c$ is \emph{any} function such that $c(s)>0, \forall s \in S$. Without loss of generality, we may and will assume in what follows that $\sum_{s\in S}c(s)=1$, i.e., $c$ is a probability distribution over $S$.

When the MDP has a large number of states, it is difficult to solve for $J^*$ using either the linear program \eqref{mdplp} or other full state representation methods such as value iteration or policy iteration \cite{BertB}. A practical solution is to resort to function approximation. The most widely used is the linear function approximation wherein the exact value function is approximated by a function that is obtained as a linear combination of a set of pre-selected basis functions.

>>>>

A stationary deterministic policy\footnote{For the scope of this paper, it suffices to restrict our attention to stationary deterministic policies.} (SDP), or simply a policy, is a map $u\colon S\ra A$ that specifies for each state what action to select in that state.
Given an SDP $u$, the expected total discounted reward corresponding to starting at state $s$, while choosing actions as dictated by $u$ for the states encountered, is
\begin{align}
J_u(s)\stackrel{\Delta}{=}\E[\sum_{n=0}^\infty \alpha^n g_{a_n}(s_n)|s_0=s,a_n=u(s_n)\mbox{ }\forall n\geq 0],\nn
\end{align}
where $\alpha \in (0,1)$ is the so-called discount factor and $s_{n+1} \sim p_{a_n}(s_n,\cdot)$, $n\ge 0$. 
We call $J_u(s)$ the value of state $s$ under SDP $u$, while $J_u$ (as a map from $S$ to the reals)
is called the value function underlying policy $u$.
The \emph{optimal policy} $u^*$ is one that in each state $s\in S$ achieves the best possible total expected discounted reward from that state. That is, $J^{u^*}(s) = J^*(s) \eqdef \max_{u\in U} J_u(s)$
where $U$ is the set of all SDPs and $J^*$ is coined the \emph{optimal value function}.%
\footnote{In our case an optimal (SDP) $u^*$ exists and is well defined \cite{BertB}.},

Any optimal policy $u^*$ and value function $J^*$ obey the Bellman equation (BE): for all $ s \in S$, 
\begin{subequations}\label{bell}
\begin{align}
\label{bellval}J^*(s)&=\max_{ a\in A}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big),\\
\label{bellpol}u^*(s)&= \underset{a\in A}{\argmax}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big)\,.
\end{align}
\end{subequations}
Once $J^*$ is computed, $u^*$ can be obtained via \eqref{bellpol}
for each state relatively cheaply (e.g., having the ability to sample from the next-state distribution at
a given state-action pair).
Thus computing the value function is at the heart of most solution methods to MDP.

The value functions $J^u$ or $J^*$ are elements of $\R^S$.
In what follows it will be useful for us to treat these as $n$-dimensional vectors, i.e., elements of $\R^n$, effectively identifying $\R^S$ with $\R^S$ in the natural way. Similarly we identify $\R^{nd}$ with $\R^{S\times A}$.
The following definitions will be useful later:
%The Bellman operator $T$ is defined using the model parameters of the MDP as follows:
\begin{definition}
Let $c,\rho,\chi:S \to \R_+$ be positive valued functions, where $\R_+$ denotes the set of strictly positive reals. Then for $J\in \R^n$, $a\in A$ and $s\in S$, 
define
\begin{enumerate}[(i)]
\item The Bellman operator $T\colon \R^n \ra \R^n$ as $(TJ)(s)=\max_{a \in A}\big(g_a(s)+\alpha \sum_{s'} p_a(s,s')J(s')\big).
$
\item \label{bellactval} The Bellman operator (of action values) $H: \R^n \to \R^{nd}$ for state-action values as $HJ=[ H_1 J,\cdots,H_d J]^\top\in \R^{nd},$ where $(H_a J)(s)= g_a(s)+\alpha \sum_{s'}p_a(s,s') J(s')$.
\item The weighted $L_1$-norms $\norm{\cdot}_{1,c}$ 
and 
the weighted $L_\infty$-norms  $\norm{\cdot}_{\infty,\rho}$ as $
||J||_{1,c}=\sum_{s \in S} c(s)|J(s)|, 
||J||_{\infty,\rho}=\max_{s \in S} \frac{|J(s)|}{\rho(s)}$.
\item The discounted maximal inflation of $\chi$ due to $P = (p_a)_{a\in A}$ as $\beta_{\chi}=\max_{s \in S} \frac{\underset{a \in A}{\max}\big(\alpha\sum_{s'}p_a(s,s')\chi(s')\big)}{\chi(s)}$.
\item The function $\chi:S\to\R_+$ above to be a \emph{Lyapunov} function for $P = (p_a)_{a\in A}$ if $\beta_{\chi}<1$.
\item $E$ to be the $nd\times n$ matrix given by $E=[I,\ldots,I]^\top$, i.e., $E$ is obtained by stacking $d$ identical $n\times n$ identity matrices one over the other.
\item A policy $\tu$ is said to be greedy with respect to (w.r.t.) $\tj\in \R^n$ if for any $s\in S$,
%some function $\tj:S \to \R$ 
\begin{align*}\tu(s)\eqdef\underset{a\in A}{\argmax}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')\tj(s')\big).\end{align*}
\end{enumerate}
\end{definition}
We now state without proof the most important properties of the Bellman operator(s).
The proofs are immediate from the definitions, but can also be found in \cite{BertB}.
First, we introduce some extra notation:
For $J_1,J_2\in \R^n$, we write $J_1\le J_2$ if $J_1(s)\le J_2(s)$ holds for all $s\in S$.
We use $\one \in \R^n$ to denote a vector with all entries $1$.
The maximum norm $\norm{\cdot}_{\infty}$ is defined by $ \norm{v}_{\infty} = \max_{s\in S} |v(s)|$.
\begin{lemma}\label{monotone}
$T$ is a monotone map, i.e., given $J_1,J_2 \in \R^n$ such that $J_1\leq J_2$, we have $T J_1\leq T J_2$. 
\end{lemma}
\begin{lemma}\label{shift}
Given $J\in \R^n$ and $t \in \R$, we have
\begin{align}\label{eq:shift}
T(J+t\one)=TJ+\alpha t\one.
\end{align}
\end{lemma}
\begin{lemma}\label{maxnorm}
If $T: \R^n \to \R^n$ is any operator that is monotonous and satisfies~\eqref{eq:shift} then 
$T$ is a $\max$-norm contraction operator with contraction factor $\alpha \in (0,1)$, i.e., given $J_1, J_2 \in \R^n$,
\begin{align}
||TJ_1-TJ_2||_\infty\leq \alpha ||J_1-J_2||_\infty.
\end{align}
\end{lemma}
\begin{lemma}\label{uniquesol} 
$J^*$ is a unique fixed point of $T$, i.e., $J^*=TJ^*$.
\end{lemma}
\begin{corollary}
If $J\in \R^n$ is such that $J\geq TJ$ then $J\geq TJ^2\geq \ldots \geq J^*$.
\end{corollary}
Though  \cref{monotone,shift,maxnorm} are stated for the Bellman operator $T$, the results also hold for $H$ as well.

\subsection{The Linear Programming Formulation}
As it is well known, the optimal value function $J^*$ can be obtained by solving the following linear program:
\begin{align}\label{mdplp}
\min_{J\in \R^n}\, &c^\top J\,\,\text{s.t.}\\
%\mb
 J(s)&\geq g_a(s)+\alpha\sum_{s'}p_a(s,s')J(s'),\,\,
\forall (s,a)\in S\times A,
\end{align}
where $c$ is \emph{any} vector of $\R^n$ such that $c(s)>0, \forall s \in S$. Without loss of generality, we may and will assume in what follows that $\sum_{s\in S}c(s)=1$, i.e., $c$ is a probability distribution over $S$.

When the MDP has a large number of states, it is difficult to solve for $J^*$ using either the linear program \eqref{mdplp} or other full state representation methods such as value iteration or policy iteration \cite{BertB}. A practical solution is to resort to function approximation. Linear function approximation, wherein the solution is searched in the subspace spanned by a set of pre-selected basis functions, is an attractive choice.
